Spatial variation in allometric growth of invasive lionfish has management implications

### 2019-01-30 docs/Manuscript.Rmd
Added all comments from reviewers, must proof read and send to Sean

---
title: "Spatial variation in allometric growth of invasive lionfish has management implications"
author: "Juan Carlos Villaseñor-Derbez, Sean Fitzgerald"
institution: "Bren School of Environmental Sciences and Management, University of California Santa Barbara, Santa Barbara, California, U.S."
abstract: |
  Lionfish (\textit{Pterois volitans / miles}) are an invasive species in the Western Atlantic and the Caribbean. Improving management of invasive lionfish populations requires accurate total biomass estimates, which depend on accurate estimates of allometric growth. Sedentary species like lionfish often exhibit high levels of spatial variation in life history characteristics. We review 17 published length-weight relationships for lionfish taken throughout their invasive range and found substantial regional differences in allometric growth parameters. The spatial pattern we observed is consistent with findings from other studies focusing on genetics or length-at-age. We show that the use of \textit{ex situ} parameters can result in up to a threefold under- or overestimation of total weight, but using parameters from nearby regions reduces this error. These findings can have substantial implications for management in terms of predicting effects on local ecosystems, evaluating the effectiveness of removal programs, or estimating biomass available for harvest.
keywords:
  - Lionfish
  - invasive species
  - length-weight
  - allometric growth
  - regional variations
bibliography: references.bib
output:
  pdf_document:
    keep_tex: yes
    number_sections: no
    template: peerj_template.tex
header-includes:
   - \usepackage{float}
   - \floatplacement{figure}{H}
   - \restylefloat{table}
---

@@

Lionfish (*Pterois volitans/miles* complex) are an invasive species in the Western Atlantic Ocean and Caribbean Sea, likely introduced through release of aquarium-kept organisms \citep{betancurr_2011}. They are the first invasive marine vertebrates established along the North Atlantic Caribbean coasts \citep{schofield_2009,schofield_2010,sabidoitza_2016} and their presence has been labeled as a major marine invasion because they threaten local biodiversity, spread rapidly, and are difficult to manage \citep{hixon_2016}. Lionfish have established invasive populations in coral reefs, estuaries, mangroves, hard-bottomed areas, and mesophotic reefs \citep{barbour_2010,jud_2011,muoz_2011,claydon_2012,andradibrown_2017,gress_2017}.

A substantial amount of research describes lionfish impacts throughout its invaded range. A meta-analysis by \citet{peake_2018} showed that invasive lionfish prey on at least 167 different species across the tropical and temperate Western Atlantic. Their feeding behavior and high consumption rates can reduce recruitment and population sizes of native reef-fish species, and can further endanger reef fish (\citet{green_2012,rocha_2015}; but see \citet{hackerott_2017}). For example, field experiments by \citet{albins_2008} showed that lionfish establishment led to reduced recruitment of native fishes by nearly 80% over a five-week period in the Bahamas. \citet{green_2012} reported that prey fish biomass declined by 65% over two years as lionfish biomass increased along Bahamian coral reefs. Their trophic impacts can be minimized if local lionfish biomass is controlled by culling \citep{ariasgonzalez_2011}.

@@

Outcomes of previous studies suggest lionfish are likely to exhibit spatial heterogeneity in the length-weight relationship, which we summarize in two main causes. First, culling programs are effective in reducing local adult populations largely because lionfish exhibit high levels of site fidelity and small home ranges \citep{Fishelson_1997,kochzius_2005,jud_2012,cote_2014}. It is known that fish with sedentary behavior are likely to exhibit high levels of spatial variation in important life history characteristics such as growth or natural mortality rates \citep{gunderson_2008,hutchinson_2008,wilson_2012,guan_2013}. Second, genetic analysis of lionfish suggests biological differences due to the existence of two genetically distinct invasive subpopulations between the Western Atlantic and the Caribbean \citep{betancurr_2011}. Site-specific parameters are necessary to accurately estimate biomass when allometric relationships are spatially variable, and this variability is increasingly important when estimating the potential effectiveness of lionfish culling programs \citep{barbour_2011,morris_2011,cote_2014,johnston_2015}. However, the region-wide differences in allometric growth parameters has remained unexplored for lionfish, despite the large number of site-specific studies reporting the length-weight relationship.

Here, we compare previously published length-weight relationships for lionfish populations in North Carolina, Northern and Southern Gulf of Mexico, the Southern Mexican Caribbean, Bahamas, Little Cayman, Jamaica, Bonaire, Puerto Rico, and Costa Rica \citep{barbour_2011,darling_2011,deleon_2013,fogg_2013,dahl_2014,edwards_2014,toledohernndez_2014,sandel_2015,aguilarperera_2016,sabidoitza_2016,sabidoitz_2016,chin_2016}. We also collected lionfish length and weight data in the central Mexican Caribbean and report the first allometric growth equation for this region. The objective of this paper is to describe the spatial pattern of length-weight relationships of lionfish across the Caribbean and Western Atlantic and to discuss implications of these spatial differences.

@@

review <- read.csv(here::here("data", "biometric_review.csv"), 
                  stringsAsFactors = F,
                  strip.white = T,
                  blank.lines.skip = T) %>% 
  mutate(c = log10(a),
         Gender2 = case_when(Gender == "Male" ~ "M",
                             Gender == "Female" ~ "F",
                             TRUE ~"B"))
```

We reviewed 12 published studies and obtained 17 length-weight relationships for the Western Atlantic (n = 2), Gulf of Mexico (n = 7), and Caribbean (n = 8, Table \ref{tab:all_params}, Fig \ref{fig:map}). We collected information on sampling methods, sex differentiation, location, and depth ranges from each study when available. Only two studies reported parameters for each gender \citep{aguilarperera_2016,fogg_2013}, so we assumed both genders were included in a study if gender was unspecified. Reviewed studies presented information for organisms obtained at depths between `r min(review$minD, na.rm = T)` m and `r max(review$maxD, na.rm = T)` m. Four studies explicitly stated that their organisms were sampled with pole spears \citep{dahl_2014,aguilarperera_2016,chin_2016,sabidoitz_2016}, and six studies mentioned that some of their organisms were obtained with pole spears (or other type of harpoon) but also hand-held nets or fish traps \citep{barbour_2011,fogg_2013,edwards_2014,toledohernndez_2014,sandel_2015,sabidoitza_2016}, and two studies did not specify how organisms were sampled \citep{darling_2011,deleon_2013}. \citet{fogg_2013} use spineless weight in their calculations, so their parameters likely underestimated total weight. Since no spineless to total weight conversions were available, these parameters were taken as reported. Recall that the objective is not to describe population-level variations, but rather to exemplify how the use of *ex situ* parameters influences estimates. We therefore still use parameters by \citet{fogg_2013}, and raise cautions whenever relevant.

@@

proj <- CRS("+proj=longlat +datum=NAD27")

@@

data(World)
W <- spTransform(World, proj) %>% 
  st_as_sf()

# Map with SF
fig_map <- ggplot() +
  geom_sf(data = W) +
  geom_point(data = review %>%
      mutate(N = ifelse(is.na(N), 1, N)),
    aes(x = lon, y = lat, fill = b, size = N), shape = 21) +
  theme_classic() +
  scale_fill_gradientn(colors = colorRamps::matlab.like(20)) +
  theme(legend.background = element_blank(),
        panel.grid.major = element_line(colour = "transparent")) +
  lims(x = c(-100, -60), y = c(9, 36)) +
  labs(x = "longitude", y = "latitude") +
  geom_text(aes(x = -70, y = 30), label = "Western Atlantic") +
  geom_text(aes(x = -90, y = 25), label = "Gulf of Mexico") +
  geom_text(aes(x = -75, y = 15), label = "Caribbean")

@@

We also used data from \citet{villaseorderbez_2014}, who collected organisms from 10 sampling sites along the central Mexican Caribbean coast in 2010 (Supplementary Table 1). Sampling locations included wall and carpet reefs at depths between 5.7 m and 38.1 m. All observed lionfish (n = 109) were collected using hand nets and numbered collection bottles. The use of hand nets prevented any weight loss due to bleeding and allowed better representation of small sizes by avoiding gear selectivity. Organisms were euthanized via pithing and Total Length (TL; mm) and Total Weight (TW; g) were recorded.

@@

Transforming this equation via base-10 logarithms we obtain:

\begin{equation}
\label{eq:log-alo}
log_{10}(TW) = b\times log_{10}(TL) + log_{10}(a)
\end{equation}

This can be simplified and re-written as:

\begin{equation}
\label{eq:log-alo-trans}
Y = bX + c
\end{equation}

Where $Y = log_{10}(TW)$, $X = log_{10}(TL)$, and $c = log_{10}(a)$. The coefficients ($c$ and $b$) were estimated with  an Ordinary Least Squares Regression and heteroskedastic-robust standard error correction \citep{zeileis_2004}. When $b = 3$, it is said that the organism exhibits a perfect isometric growth, so the $b$ coefficient was tested against the null hypothesis of isometric growth (*i.e.* $H_0: b = 3$). Coefficients were tested with a two-tailed Student's t, and the significance of the regression was corroborated with an F-test.

Some of the reviewed studies inconsistently defined $a$ as either the ponderal index from Eq. \ref{eq:allometric} or the y-intercept ($c$) from Eq. \ref{eq:log-alo-trans}. Other studies incorrectly reported parameters as mm-to-g conversions when they were in fact cm-to-g conversions. We standardized each study by converting coefficients and report all parameters as TL(mm) to TW (g) conversions. Locations where allometric studies have been performed are shown in Figure \ref{fig:map} and summarized in Table \ref{tab:all_params}.

We obtained a total of 18 parameter pairs by combining length-weight parameters extracted from the literature and the additional pair calculated here. We used the central Mexican Caribbean as a case study of how the use of *ex situ* parameters influences the accuracy of weight estimates for lionfish. We estimated TW from the TL observations we collected in the central Mexican Caribbean (n = 109, with $TL$ ranging from 34 to 310 mm). using each of the 18 parameter pairs and divided predicted weights by known observed weights to obtain a simple measure of over- or underestimation. Difference in mean weight ratios across the different parameter pairs were tested with a one-way analysis of variance (ANOVA) and Tukey's test was used for *post-hoc* tests. All analyses were performed in R version `r paste(version$major, version$minor, sep = ".")` \citep{rcore_2018}. Raw data and code used in this work are available on github at [https://github.com/jcvdav/lionfish_biometry](https://github.com/jcvdav/lionfish_biometry).

@@

```{r}
Predicted_biomass <- review %>% 
  mutate(Length = list(fish$Length),
         Weight = list(fish$Weight)) %>% 
  unnest() %>% 
  mutate(pWeight = a*(Length^b),
    Biomass_ratio = pWeight / Weight)

anova <- aov(Biomass_ratio ~ Study_id, data = Predicted_biomass)
ancova_ratios <- aov(car::logit(Biomass_ratio) ~ Study_id + Length, data = Predicted_biomass)

ancova <- aov(formula = log10(Weight / max(Weight)) ~ Study_id * log10(Length / max(Length)), data = Predicted_biomass)

groups <- HSD.test(y = anova, trt = "Study_id")

@@

The length-weight relationship for organisms from the central Mexican Caribbean resulted in the coefficient values $a = `r 10^c`$, $b = `r b`$ and $c = `r c`$ ($R^2 = `r Rsq`$, `r Fstring`, $p < 0.001$). The allometric factor ($b$) was significantly different from $b = 3$ ($t(107) = `r formatC((coef(model)[[2]]-3)/robust_se[[2]], digits = 2, format = "f")`; p<0.001$) indicating that lionfish present allometric growth. The length-weight coefficients estimated in this study were within the range identified by studies in other regions (Table \ref{tab:all_params}). Figure \ref{fig:l-w-carib} shows the relationship between TL and TW for this region, and model fit statistics are presented in Table \ref{tab:reg_table}.

@@

```{r, results = "asis", eval = F}
stargazer(model,
          dep.var.caption = "$log_{10}(TW)$",
          dep.var.labels.include = F,
          header = F,
          se = list(robust_se),
          type = "latex",
          t.auto = T,
          p.auto = T,
          intercept.bottom = F,
          single.row = T,
          report = "vcs*",
          omit.stat = c("f", "rsq"),
          add.lines = list(f_string),
          covariate.labels = c("c", "b"),
          dep.var.labels = "log10(TW)", title = "\\label{tab:reg_table}Coefficients of the linear model fit to Eq \\ref{eq:log-alo-trans}. Numbers in parenthesees represent heteroskedastic-robust standard errors.")
```

There were significant differences in our predicted weights for the central Mexican Caribbean when using each of the different pairs of parameters ($F(df = 17; 1944) = 61.55; p < 0.001$). The lowest weight estimates for the observed lengths resulted from using the allometric parameters from Banco Chinchorro in the Caribbean, with mean $\pm$ SD of `r filter(Predicted_biomass, Study_id == "Sabido2BC") %$% mean(pWeight) %>% formatC(digits = 2, format = "f")` $\pm$ `r filter(Predicted_biomass, Study_id == "Sabido2BC") %$% sd(pWeight) %>% formatC(digits = 2, format = "f")` g \citep{sabidoitz_2016}. In contrast, the highest weight estimates came from the Western Atlantic with `r filter(Predicted_biomass, Study_id == "Barbour") %$% mean(pWeight) %>% formatC(digits = 2, format = "f")` $\pm$ `r filter(Predicted_biomass, Study_id == "Barbour") %$% sd(pWeight) %>% formatC(digits = 2, format = "f")` g \citep{barbour_2011}. To put this in context, true observed weights have a mean of `r filter(Predicted_biomass, Study_id == "Barbour") %$% mean(Weight) %>% formatC(digits = 2, format = "f")` $\pm$ `r filter(Predicted_biomass, Study_id == "Barbour") %$% sd(Weight) %>% formatC(digits = 2, format = "f")` g. Weigths predicted from these extreme parameters correspond to predicted-to-observed weights ratios of `r filter(Predicted_biomass, Study_id == "Sabido2BC") %$% mean(Biomass_ratio) %>% formatC(digits = 2, format = "f")` $\pm$ `r filter(Predicted_biomass, Study_id == "Sabido2BC") %$% sd(Biomass_ratio) %>% formatC(digits = 2, format = "f")` and `r filter(Predicted_biomass, Study_id == "Barbour") %$% mean(Biomass_ratio) %>% formatC(digits = 2, format = "f")` $\pm$ `r filter(Predicted_biomass, Study_id == "Barbour") %$% sd(Biomass_ratio) %>% formatC(digits = 2, format = "f")` (mean $\pm$ SD), respectively.

The calculated ratio of predicted-to-observed weight ranged from `r min(Predicted_biomass$Biomass_ratio) %>% formatC(digits = 2, format = "f")` to `r max(Predicted_biomass$Biomass_ratio) %>% formatC(digits = 2, format = "f")`, indicating that *ex situ* parameters can result in substantial weight under- and overestimation. Tukey's *post-hoc* test suggests that weight ratios for the central Mexican Caribbean were not different from those obtained with parameters from Little Cayman, the Bahamas, and some sites in the Gulf of Mexico (Tukey's HSD $p > 0.05$). Weight estimates using parameters from the Gulf of Mexico and Western Atlantic were higher on average than those from the Caribbean (Fig \ref{fig:all_allo}). The average ($\pm$ SD) predicted-to-observed weight ratios from these three regions were 1.24 $\pm$ 0.309, 1.76 $\pm$ 0.496, and 1.17 $\pm$ 0.398, respectively. Predicted-to-observed weight ratios are presented in Figure \ref{fig:bio_ratio}. Spineless weight parameters from \citet{fogg_2013} still produced predicted-to-observed weight ratios > 1.

```{r}
review %>% 
  select(Region, Gender2, N, a, b, r2, Study) %>% 
  filter(!Study == "Fishbase") %>% 
  arrange(desc(Region)) %>% 
  mutate(r2 = ifelse(is.na(r2), "-", r2),
         N = ifelse(is.na(N), "-", N),
         a = formatC((a*1e5), digits = 2)) %>% 
  knitr::kable(row.names = F,
               digits =  c(0, 0, 0, 0, 2, 2),
               booktabs = T,
               col.names = c("Region", "Sex", "n", "a", "b", "$R^2$", "Reference"),
               caption = "\\label{tab:all_params}Summary of 18 allometric growth parameters available for lionfish in the invaded range from peer-reviewed literature and this study. All parameters have been adjusted to convert from millimeters to grams. n = Sample size, Sex specifies whether data was presented for Females (F), Males (M), or both genders combined (B), a = scaling parameter for Eq. 1 (presented in $\\times 10^{-5}$), b = exponent or slope for Eq. 1 or Eq. 3, respectively.",
               escape = T)
```

```{r}
fig_all_allo <- review %>% 
  filter(!Study == "This study") %>% 
  mutate(Length = list(seq(1, 400))) %>% 
  unnest() %>% 
  mutate(Weight = a*(Length^b),
         This = 10^(coefficients(model)[1] + coefficients(model)[2]*log10(Length))) %>%
  group_by(Length) %>% 
  mutate(min = min(Weight),
         max = max(Weight)) %>% 
  ungroup() %>% 
  ggplot(aes(x = Length)) +
  geom_line(aes(y = Weight, color = Study, linetype = Gender, group = Study_id), size = 1) +
  geom_line(aes(y = This), size = 1.2, color = "black", linetype = "dashed") +
  labs(x = "Total length (mm)", y = "Total Weight (g)") +
  theme_cowplot() +
  theme(text = element_text(family = "Times"),
        legend.justification=c(0,1),
        legend.position=c(0.01, 1.05),
        legend.background = element_rect(fill = "transparent"),
        legend.box = "horizontal",
        legend.text = element_text(size = 9),
        legend.title = element_text(size = 10)) +
  guides(color = guide_legend(order = 1),
         linetype = guide_legend(order = 0, ncol = 2, override.aes = list(size = 0.5)),
         size = "none") +
  scale_color_brewer(palette = "Paired", direction = -1) +
  scale_linetype_manual(values=c("solid", "dashed", "dotted"), name = "Sex") +
  scale_size_manual(values = c(1, 1.5))
```

```{r fit2, fig.height = 3.8, fig.pos = "H", fig.cap = "\\label{fig:all_allo}Length-weight relationships (n = 18) for 12 studies and this study. Colors indicate studies from which the parameters were extracted. Dotted, dashed and solid lines show models for males, females, and combined sexes, respectively. The dashed black line represents the relationship estimated in this study. There are two solid green lines for Sabido-Itza et al, 2016b, one for each of the two sites for which they report parameters."}
fig_all_allo
```

>>>>

---
title: "Spatial variation in allometric growth of invasive lionfish has management implications"
author: "Juan Carlos Villaseñor-Derbez, Sean Fitzgerald"
institution: "Bren School of Environmental Sciences and Management, University of California Santa Barbara, Santa Barbara, California, U.S."
abstract: |
  Lionfish (\textit{Pterois volitans / miles}) are an invasive species in the Western Atlantic and the Caribbean. Improving management of invasive lionfish populations requires accurate total biomass estimates, which depend on accurate estimates of allometric growth. Sedentary species like lionfish often exhibit high levels of spatial variation in life history characteristics. We review 17 published length-weight relationships for lionfish taken throughout their invasive range and found significant regional differences in allometric growth parameters. The spatial pattern we observed is consistent with findings from other studies focusing on genetics or length-at-age. We show that the use of \textit{ex situ} parameters can result in up to a threefold under- or overestimation of total weight, but using parameters from nearby regions reduces this error. These findings can have substantial implications for management in terms of predicting effects on local ecosystems, evaluating the effectiveness of removal programs, or estimating biomass available for harvest.
keywords:
  - Lionfish
  - invasive species
  - length-weight
  - allometric growth
  - regional variations
bibliography: references.bib
output:
  pdf_document:
    keep_tex: yes
    number_sections: no
    template: peerj_template.tex
header-includes:
   - \usepackage{float}
   - \floatplacement{figure}{H}
   - \restylefloat{table}
---

@@

Lionfish (*Pterois volitans/miles* complex) are an invasive species in the Western Atlantic Ocean and Caribbean Sea, likely introduced through release of aquarium-kept organisms \citep{betancurr_2011}. They are the first invasive marine vertebrates established along these coasts \citep{schofield_2009,schofield_2010,sabidoitza_2016} and their presence has been labeled as a major marine invasion because they threaten local biodiversity, spread rapidly, and are difficult to manage \citep{hixon_2016}. Lionfish have established invasive populations in coral reefs, estuaries, mangroves, hard-bottomed areas, and mesophotic reefs \citep{barbour_2010,jud_2011,muoz_2011,claydon_2012,andradibrown_2017,gress_2017}.

A substantial amount of research describes lionfish impacts throughout its invaded range. A meta-analysis by \citet{peake_2018} showed that invasive lionfish prey on at least 167 different species across the tropical and temperate Western Atlantic. Their feeding behavior and high consumption rates can reduce recruitment and population sizes of native reef-fish species, and can further endanger reef fish \citep{green_2012,rocha_2015}; but see \citet{hackerott_2017}. For example, field experiments by \citet{albins_2008} showed that lionfish establishment led to reduced recruitment of native fishes by nearly 80% over a five-week period in the Bahamas. \citet{green_2012} reported that prey fish biomass declined by 65% over two years as lionfish biomass increased along Bahamian coral reefs. However, their trophic impacts can be minimized if local lionfish biomass is controlled by culling \citep{ariasgonzalez_2011}.

@@

Outcomes of previous studies suggest lionfish are likely to exhibit spatial heterogeneity in the length-weight relationship, which we summarize in two main causes. First, culling programs are effective in reducing local adult populations largely because lionfish exhibit high levels of site fidelity and small home ranges \citep{Fishelson_1997,kochzius_2005,jud_2012,cote_2014}. It is known that fish with sedentary behavior are likely to exhibit high levels of spatial variation in important life history characteristics such as growth or natural mortality rates \citep{gunderson_2008,hutchinson_2008,wilson_2012,guan_2013}. Second, genetic analysis of lionfish suggests biological differences due to the existence of two genetically distinct invasive subpopulations between the Western Atlantic and the Caribbean \citep{betancurr_2011}. The large number of site-specific studies reporting the length-weight relationship of lionfish provide variable estimates. These differences may be increasingly important when estimating the potential effectiveness of lionfish culling programs \citep{barbour_2011,morris_2011,cote_2014,johnston_2015}. However, the extent to which using *ex situ* parameters to estimate total weight from length observations remains unexplored.

Here, we use previously published length-weight relationships for lionfish populations in North Carolina, Northern and Southern Gulf of Mexico, the Southern Mexican Caribbean, Bahamas, Little Cayman, Jamaica, Bonaire, Puerto Rico, and Costa Rica to quantify the magnitude of the error caused by using *ex situ* parameters to estimate lionfish weight from length observations. We also collected lionfish length and weight data in the central Mexican Caribbean and report the first length-weight relationship for this region.

<!-- \citep{barbour_2011,darling_2011,deleon_2013,fogg_2013,dahl_2014,edwards_2014,toledohernndez_2014,sandel_2015,aguilarperera_2016,sabidoitza_2016,sabidoitz_2016,chin_2016} -->

@@

review <- read.csv(here::here("data", "biometric_review.csv"), 
                  stringsAsFactors = F,
                  strip.white = T,
                  blank.lines.skip = T) %>% 
  mutate(c = log10(a),
         Gender2 = case_when(Gender == "Male" ~ "M",
                             Gender == "Female" ~ "F",
                             TRUE ~"B"),
         tl_min = ifelse(is.na(tl_min), 34, tl_min),
         tl_max = ifelse(is.na(tl_max), 310, tl_max))
```

We reviewed 12 published studies and obtained 17 length-weight relationships for the Western Atlantic (n = 2), Gulf of Mexico (n = 7), and Caribbean (n = 8, Table \ref{tab:all_params}, Fig \ref{fig:map}). We collected information on sampling methods, sex differentiation, location, and depth ranges from each study when available. Only two studies reported parameters for each gender \citep{aguilarperera_2016,fogg_2013}, so we assumed both genders were included in a study if gender was unspecified. Reviewed studies presented information for organisms obtained at depths between `r min(review$minD, na.rm = T)` m and `r max(review$maxD, na.rm = T)` m. Four studies explicitly stated that their organisms were sampled with pole spears \citep{dahl_2014,aguilarperera_2016,chin_2016,sabidoitz_2016}, and six studies mentioned that some of their organisms were obtained with pole spears (or other type of harpoon) but also hand-held nets or fish traps \citep{barbour_2011,fogg_2013,edwards_2014,toledohernndez_2014,sandel_2015,sabidoitza_2016}, and two studies did not specify how organisms were sampled \citep{darling_2011,deleon_2013}.

@@

proj <- c("+proj=longlat +datum=NAD27")

@@

data(World)
W <- st_transform(World, proj)

# Map with SF
fig_map <- ggplot() +
  geom_sf(data = W) +
  geom_point(data = review %>%
      mutate(N = ifelse(is.na(N), 1, N)),
    aes(x = lon, y = lat, fill = b, size = N), shape = 21) +
  theme_classic() +
  scale_fill_gradientn(colors = colorRamps::matlab.like(20)) +
  theme(legend.background = element_blank(),
        panel.grid.major = element_line(colour = "transparent")) +
  lims(x = c(-100, -60), y = c(9, 36)) +
  labs(x = "Longitude", y = "Latitude") +
  geom_text(aes(x = -70, y = 30), label = "Western Atlantic") +
  geom_text(aes(x = -90, y = 25), label = "Gulf of Mexico") +
  geom_text(aes(x = -75, y = 15), label = "Caribbean")

@@

We also used data from \citet{villaseorderbez_2014}, who collected organisms from 10 sampling sites along the central Mexican Caribbean coast in the Summer of 2010 (Supplementary Table 1). Sampling locations included wall and carpet reefs at depths between 5.7 m and 38.1 m. All observed lionfish (n = 109) were collected using hand nets and numbered collection bottles. The use of hand nets prevented any weight loss due to bleeding and allowed better representation of small sizes by avoiding gear selectivity. Organisms were euthanized via pithing and Total Length (TL; mm) and Total Weight (TW; g) were recorded.

@@

<!-- Transforming this equation via base-10 logarithms we obtain: -->

<!-- \begin{equation} -->
<!-- \label{eq:log-alo} -->
<!-- log_{10}(TW) = b\times log_{10}(TL) + log_{10}(a) -->
<!-- \end{equation} -->

<!-- This can be simplified and re-written as: -->

<!-- \begin{equation} -->
<!-- \label{eq:log-alo-trans} -->
<!-- Y = bX + c -->
<!-- \end{equation} -->

<!-- Where $Y = log_{10}(TW)$, $X = log_{10}(TL)$, and $c = log_{10}(a)$. -->

The above equation was linearized using $log_{10}$-transformation. The coefficients were estimated with an Ordinary Least Squares Regression, and heteroskedastic-robust standard error correction was applied \citep{zeileis_2004}. When $b = 3$, it is said that the organism exhibits a perfect isometric growth, so the $b$ coefficient was tested against the null hypothesis of isometric growth (*i.e.* $H_0: b = 3$). Coefficients were tested with a two-tailed Student's t, and the significance of the regression was corroborated with an F-test.

Some of the reviewed studies inconsistently defined $a$ as either the ponderal index from Eq. \ref{eq:allometric} or the y-intercept from the linearized log-transformed equation. Other studies incorrectly reported parameters as mm-to-g conversions when they were in fact cm-to-g conversions. We standardized each study by converting coefficients and report all parameters as TL(mm) to TW (g) conversions. Locations where allometric studies have been performed are shown in Figure \ref{fig:map} and summarized in Table \ref{tab:all_params}.

We obtained a total of 18 parameter pairs by combining length-weight parameters extracted from the literature and the additional pair calculated here. Recall that the objective of this study is not to describe population-level variations, but rather to estimate how the use of *ex situ* parameters influences weight estimates. We used the central Mexican Caribbean as a case study of how the use of *ex situ* parameters influences the accuracy of weight estimates for lionfish. We estimated $TW$ from the $TL$ observations we collected in the central Mexican Caribbean (n = 109, with $TL$ ranging from 34 mm to 310 mm) using each of the 18 parameter pairs and divided predicted weights by known observed weights to obtain a simple measure of over- or underestimation. Difference in mean weight ratios were tested with an analysis of covariance (ANCOVA) using study and $TL$ as covariates. Ratios were logit-transformed prior to analysis, and a *post-hoc* Tukey's test was used to identify groups where mean ratios did not differ. All analyses were performed in R version `r paste(version$major, version$minor, sep = ".")` \citep{rcore_2018}. Raw data and code used in this work are available on github at github.com/jcvdav/lionfish_biometry.

@@

```{r}
Predicted_biomass <- review %>% 
  mutate(Length = list(fish$Length),
         Weight = list(fish$Weight)) %>% 
  unnest() %>% 
  mutate(pWeight = a*(Length^b),
    Biomass_ratio = pWeight / Weight,
    logit_biomass_ratio = car::logit(Biomass_ratio))

anova <- aov(Biomass_ratio ~ Study_id, data = Predicted_biomass)
ancova <- aov(logit_biomass_ratio ~ Study_id + Length, data = Predicted_biomass)

groups <- HSD.test(y = ancova, trt = "Study_id")

@@

The length-weight relationship for organisms from the central Mexican Caribbean resulted in the coefficient values $a = `r 10^c`$, $b = `r b`$ and $c = `r c`$ ($R^2 = `r Rsq`$, `r Fstring`, $p < 0.001$). The allometric factor ($b$) was significantly different from $b = 3$ ($t(107) = `r formatC((coef(model)[[2]]-3)/robust_se[[2]], digits = 2, format = "f")`; p<0.001$) indicating that lionfish present allometric growth. The length-weight coefficients estimated in this study were within the range identified by studies in other regions (Table \ref{tab:all_params}). Figure \ref{fig:l-w-carib} shows the relationship between TL and TW for this region.

@@

```{r, results = "asis", eval = F}
stargazer(model,
          dep.var.caption = "$log_{10}(TW)$",
          dep.var.labels.include = F,
          header = F,
          se = list(robust_se),
          type = "latex",
          t.auto = T,
          p.auto = T,
          intercept.bottom = F,
          single.row = T,
          report = "vcs*",
          omit.stat = c("f", "rsq"),
          add.lines = list(f_string),
          covariate.labels = c("c", "b"),
          dep.var.labels = "log10(TW)", title = "\\label{tab:reg_table}Coefficients of the linear model fit to Eq 1. Numbers in parenthesees represent heteroskedastic-robust standard errors.")
```

There were significant differences in our predicted weights for the central Mexican Caribbean when using each of the different pairs of parameters ($F(df = 17; 1944) = 61.55; p < 0.001$). The lowest weight estimates for the observed lengths resulted from using the allometric parameters from Banco Chinchorro in the Caribbean, with mean $\pm$ SD of `r filter(Predicted_biomass, Study_id == "Sabido2BC") %$% mean(pWeight) %>% formatC(digits = 2, format = "f")` $\pm$ `r filter(Predicted_biomass, Study_id == "Sabido2BC") %$% sd(pWeight) %>% formatC(digits = 2, format = "f")` g \citep{sabidoitz_2016}. In contrast, the highest weight estimates came from the Western Atlantic with `r filter(Predicted_biomass, Study_id == "Barbour") %$% mean(pWeight) %>% formatC(digits = 2, format = "f")` $\pm$ `r filter(Predicted_biomass, Study_id == "Barbour") %$% sd(pWeight) %>% formatC(digits = 2, format = "f")` g \citep{barbour_2011}. To put this in context, true observed weights have a mean of `r filter(Predicted_biomass, Study_id == "Barbour") %$% mean(Weight) %>% formatC(digits = 2, format = "f")` $\pm$ `r filter(Predicted_biomass, Study_id == "Barbour") %$% sd(Weight) %>% formatC(digits = 2, format = "f")` g. Weigths predicted from these extreme parameters correspond to mean $\pm$ SD predicted-to-observed weights ratios of `r filter(Predicted_biomass, Study_id == "Sabido2BC") %$% mean(Biomass_ratio) %>% formatC(digits = 2, format = "f")` $\pm$ `r filter(Predicted_biomass, Study_id == "Sabido2BC") %$% sd(Biomass_ratio) %>% formatC(digits = 2, format = "f")` and `r filter(Predicted_biomass, Study_id == "Barbour") %$% mean(Biomass_ratio) %>% formatC(digits = 2, format = "f")` $\pm$ `r filter(Predicted_biomass, Study_id == "Barbour") %$% sd(Biomass_ratio) %>% formatC(digits = 2, format = "f")` (mean $\pm$ SD), respectively. At its extremes, other parameters resulted in under- and overestimations in the order of `r min(Predicted_biomass$Biomass_ratio) %>% formatC(digits = 2, format = "f")` to `r max(Predicted_biomass$Biomass_ratio) %>% formatC(digits = 2, format = "f")` of the actual observed weight, indicating that *ex situ* parameters can result in substantial weight under- and overestimation.

Tukey's *post-hoc* test suggests that weight ratios for the central Mexican Caribbean were not different from those obtained with parameters from Little Cayman, the Bahamas, and some sites in the Gulf of Mexico (Tukey's HSD $p > 0.05$). Weight estimates using parameters from the Gulf of Mexico and Western Atlantic were higher on average than those from the Caribbean (Fig \ref{fig:all_allo}). The average ($\pm$ SD) predicted-to-observed weight ratios from these three regions were 1.24 $\pm$ 0.309, 1.76 $\pm$ 0.496, and 1.17 $\pm$ 0.398, respectively. Predicted-to-observed weight ratios are presented in Figure \ref{fig:bio_ratio}.

```{r}
review %>% 
  select(Region, Gender2, N, a, b, r2, Study) %>% 
  filter(!Study == "Fishbase") %>% 
  arrange(desc(Region)) %>% 
  mutate(r2 = formatC(r2, digits = 2),
         r2 = ifelse(r2 == " NA", "-", r2),
         N = ifelse(is.na(N), "-", N),
         a = formatC((a*1e5), digits = 2)) %>% 
  knitr::kable(row.names = F,
               digits =  c(0, 0, 0, 0, 2, 2),
               booktabs = T,
               col.names = c("Region", "Sex", "n", "a", "b", "$R^2$", "Reference"),
               caption = "\\label{tab:all_params}Summary of 18 allometric growth parameters available for lionfish in the invaded range from peer-reviewed literature and this study. All parameters have been adjusted to convert from millimeters to grams. n = Sample size, Sex specifies whether data was presented for Females (F), Males (M), or both genders combined (B), a = scaling parameter (presented in $\\times 10^{-5}$), b = exponent.")
```

```{r}
This <- data.frame(Length = c(34:310)) %>% 
  mutate(Weight = 10^(coefficients(model)[1] + coefficients(model)[2]*log10(Length)))

fig_all_allo <- review %>% 
  filter(!Study == "This study") %>%
  mutate(Length = list(seq(1, 400))) %>% 
  unnest() %>% 
  filter(Length > tl_min & Length < tl_max) %>% 
  mutate(Weight = a*(Length^b)) %>%
  group_by(Length) %>% 
  mutate(min = min(Weight),
         max = max(Weight)) %>% 
  ungroup() %>% 
  ggplot(aes(x = Length)) +
  geom_line(aes(y = Weight, color = Study, linetype = Gender, group = Study_id), size = 1) +
  geom_line(data = This, aes(x = Length, y = Weight),
            size = 1.2, color = "black", linetype = "dashed") +
  labs(x = "Total length (mm)", y = "Total Weight (g)") +
  theme_cowplot() +
  theme(text = element_text(family = "Times"),
        legend.justification=c(0,1),
        legend.position=c(0.01, 1.05),
        legend.background = element_rect(fill = "transparent"),
        legend.box = "horizontal",
        legend.text = element_text(size = 9),
        legend.title = element_text(size = 10)) +
  guides(color = guide_legend(order = 1),
         linetype = guide_legend(order = 0, ncol = 2, override.aes = list(size = 0.5)),
         size = "none") +
  scale_color_brewer(palette = "Paired", direction = -1) +
  scale_linetype_manual(values=c("solid", "dashed", "dotted"), name = "Sex") +
  scale_size_manual(values = c(1, 1.5))
```
